---
title: "Assesssment 1 - COVID-19 impact on digital learning"
author: Provide Student Name and Student ID
output: word_document
date: "2024-03-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## 1. Data Cleaning and Wrangling  
**Describe and show the data cleaning, transformation and wrangling process' that you have completed. Ensure you provide justification of any data handling and transformation you have completed.**


<br>
<br>

you can embed your transfomration code and provide justification like this:  

example: This code creates x and y variable, i have created these variable so that i can add these to my data frame and compute the required measures. Please remember to use comments in your code.

```{r, summary-stats, echo=TRUE}
x <- 1:10    # create an x variable
y <- 10:1    # create a y variable
dataf <- data.frame(x = x, y = y)

summary(dataf)
```
  
\newpage


## 2. Data Visualisation
**Provide evidence of your visualisation techniques and the reason for your choice of visual tools. Provide a description of each visualisation you are presenting. There is no need to show your code.**

<br>

You can also embed plots, and provide reason for your choice like this:  

example: I have used the below visualisation technique because .... 

This visualisation shows the following...


```{r, simple-plot, echo=FALSE}
x <- 1:10    # create an x variable
y <- 10:1    # create a y variable
dataf <- data.frame(x = x, y = y)
plot(dataf$x, dataf$y, xlab = "x axis", ylab = "y axis")
```

\newpage


## 3. Findings  
**Clearly explain your findings and provide a valid conclusion of your analysis. References may be used if required.(max 500 words) **







<!-- Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot. -->